%%% Seznam použité literatury je zpracován podle platných standardù. Povinnou citaèní
%%% normou pro bakaláøskou práci je ISO 690. Jména èasopisù lze uvádìt zkrácenì, ale jen
%%% v kodifikované podobì. Všechny použité zdroje a prameny musí být dùslednì citovány.

\def\bibname{Bibliography}
\begin{thebibliography}{99}
\addcontentsline{toc}{chapter}{\bibname}

\bibitem{temko}
	{\sc TEMKO, }Florence; {\sc JACKSON, }Paul.
	\emph{Paper pandas and jumping frogs}.
	China : China books \& periodicals, 1986. 133 p. 
	ISBN 0-8351-17701770-7.

\bibitem{lang}
	{\sc LANG, }Robert J.
	\emph{Origami design secrets : mathematical methods for an ancient art}.
	Massachusetts (USA) : A K Peters, Ltd., 2003. 585 p.
	ISBN 1-56881-194-2

\bibitem{pen}
	{\sc Pen, }K. 
	\emph{Origami Paper Folding: Rules of the Game}.
	Associated content [online]. 
	2009 [cited 2011-05-25].
	Available at WWW: \url{http://www.associatedcontent.com/article/1818997/origami_paper_folding_rules_of_the.html?cat=24}.

\bibitem{fuse}
	{\sc Fuse, }Tomoko.
	\emph{Unit Origami: Multidimensional Transformations}. 
	Japan : Japan Publications, 1990.
	ISBN 0870408526.

\bibitem{lang2}
	{\sc Lang, }Robert J.
	\emph{Origami in Action: Paper Toys that Fly, Flap, Gobble and Inflate}.
	Massachusetts (USA) : St. Martin's Griifin, 1997.
	ISBN 0-312-15618-9.

\bibitem{smith}
	{\sc Smith, }John S.
	\emph{Pureland Origami}. 
	Great Britain: British Origami Society, 1980.

\bibitem{temko2}
	{\sc TEMKO, }Florence.
	\emph{Kirigami Home Decorations}.
	China : Tuttle Publishing, 2006.
	ISBN 0-8048-3793-7.

\bibitem{mobilereference}
	{\sc MobileReference}.
	\emph{Asian Art}.
	USA : SoundTells, 2003.

\bibitem{folds}
	{\sc Wikipedia contributors}.
	\emph{Origami techniques}.
	Wikipedia, The Free Encyclopedia [online]. 
	2011 [cited 2011-05-25].
	Available at WWW: \url{http://en.wikipedia.org/w/index.php?title=Origami_techniques&oldid=420452758}

\bibitem{demaine}
	{\sc Demaine, }E. D.; {\sc Demaine, }M. L.
	\emph{Recent results in computational origami.}
	In Origami : Proceedings of the 3rd International Meeting of Origami Science, Math, and Education. 
	California (USA) : Monterey, 2001. pp. 3--16. Available at WWW: \url{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.100.8809&rep=rep1&type=pdf}

\bibitem{bern}
	{\sc Bern, }Marshall; {\sc Hayes, }Barry.
	\emph{The complexity of flat origami}.
	In Proceedings of the 7th Annual ACM-SIAM Symposium on Discrete Algorithms, pages 175ï¿½183, Atlanta, January 1996.
	Reprinted in Proceedings of the 3rd International Meeting of Origami Science, Math, and Education, 2001

\bibitem{treemaker}
	{\sc Lang, }Robert J.
	\emph{TreeMaker 4.0:  A Program for Origami Design}. [online]
	1998 [cited 2011-05-25].
	Available at WWW: \url{http://www.langorigami.com/science/treemaker/TreeMkr40.pdf}

\bibitem{simulation}
	{\sc Miyazaki, }Shin-ya; {\sc Yasuda, }Takami; {\sc Yokoi, }Shigeki; {\sc Toriwaki, }Jun-ichiro.
	\emph{An ORIGAMI Playing Simulator in the Virtual Space}.
	In The Journal of Visualization and Computer Animation, vol.7(1)
	Japan, 1996. pp. 25--42. Available at WWW: \url{http://www.om.sist.chukyo-u.ac.jp/main/research/origami/journal/jvca.html}

\bibitem{eos}
	{\sc Ida, }Tetsuo; {\sc Takahashi, }Hidekazu; {\sc Marin,
}Mircea; {\sc Kasem, }Asem; {\sc Ghourabi, }Fadoua.
	\emph{Computational Origami System Eos}
	Japan : Department of Computer Science, University of Tsukuba. 
	Available at WWW: \url{http://www2.score.cs.tsukuba.ac.jp/publications/Eos.pdf/view}

\bibitem{stent}
	{\sc You, }Zhong; {\sc Kuribayashi, }Kaori
	\emph{A NOVEL ORIGAMI STENT}
	In 2003 Summer Bioengineering Conferencee, June 25--29.
	Florida (USA), 2003. pp. 257--258.
	Available at WWW: \url{http://www.tulane.edu/~sbc2003/pdfdocs/0257.PDF}

\bibitem{eyeglass}
	{\sc Lawrence Livermore National Laboratory}.
	\emph{Eyeglass space telescope}.
	National Ignition Facility \& Photon Science : Bringing Start power to Earth [online].
	2002 [cited 2011-05-25]. 
	Available at WWW: \url{https://lasers.llnl.gov/programs/psa/pdfs/technologies/eyeglass_space_telescope.pdf}

\bibitem{sfu}
	{\sc Shibayama}, Y.; {\sc Arai}, H.; {\sc Matsui}, K.; {\sc Hama}, K.; {\sc Ushirokawa}, A.; {\sc Natori}, M.; {\sc Takahashi}, K.; {\sc Wakasugi}, N.; {\sc Anzai}, T.
	\emph{SFU solar array}.
	European Space Power Conference. ESA Special Publication. 1989. , pp. 557-562.
	Available at WWW: \url{http://adsabs.harvard.edu/abs/1989ESASP.294..557S}

\bibitem{airbag}
	{\sc Gï¿½rtner, }Torsten; {\sc Eriksson, }Magnus; {\sc Fï¿½ltstrï¿½m, }Jonas.
	EASi GmbH, \emph{Advanced Technologies for the Simulation of Folded Airbags}, 2nd European LS-DYNA Conference, Gothenburg, Sweden, June 1999.

\bibitem{threedpanels}
	{\sc Choi, }Charles Q.
	\emph{Solar Panel Productivity Boosted by Origami}.
	LiveScience.com [online]. 2010 [cited 2011-05-25].
	Available at WWW: \url{http://www.livescience.com/10625-solar-panel-productivity-boosted-origami.html}

\bibitem{javasupport}
	{\sc Oracle}
	\emph{Supported System Configurations for Java SE 6 and Java For Business 6}
	Oracle.com [online]. [cited 2011-05-25].
	Available at WWW: \url{http://www.oracle.com/technetwork/java/javase/system-configurations-135212.html}

\bibitem{newjava}
	{\sc Oracle}
	\emph{Next Generation in Applet Java Plug-in Technology}
	Oracle : Sun Developer Network [online]. [cited 2011-05-25].
	Available at WWW: \url{http://java.sun.com/developer/technicalArticles/javase/newapplets/}

\bibitem{java3dplatforms}
	\emph{Java 3D TM Downloads: Release Builds}
	java.net : The Source for Java Technology Collaboration [online]. [cited 2011-05-25].
	Available at WWW: 
	\url{http://java3d.java.net/binary-builds.html}

\bibitem{dawson}
	{\sc Dawson, }Bruce.
	\emph{Comparing floating point numbers}
	Cygnus-software.com [online]. [cited 2011-05-26].
	Available at WWW: \url{http://www.cygnus-software.com/papers/comparingfloats/Comparing\%20floating\%20point\%20numbers.htm}

\end{thebibliography}
